NMRlipids IV: Headgroup \& glycerol backbone structures, and cation binding in bilayers with PS lipids

### 2018-12-07 Manuscript/manuscriptPS.tex
New version of conclusions

\section{Conclusions}
We have collected a set of experimental NMR order parameter data,
which could be combined with MD simulations to interpret the
headgroup structure and cation binding details to negatively charged
membranes containing PS lipids. Using open collaboration method, we
tried to find a MD simulation model which would be sufficiently accurate
to interpret the experimental data. However, none of the tested models
was accurate enough. In line with the previous study for PC lipids \cite{catte16},
MD simulation models seems to generally overestimate cation binding also
to negatively charged bilayers containing PS lipids, with some exceptions.
The response of PS lipid headgroup order parameters to the bound cations
does not agree with experiments, even in the cases where binding affinity is
not overestimated. This is in contrast to the previous results with PC lipids,
where the qualitative response of the headgroup order parameters was in agreement
with experiments even in the cases where the headgroup structure without ions was not
correct and the cation binding affinity was overestimated. In addition, the inaccurate
responses of PS headgroup order parameters to the dilution with PC lipids suggests
that the PC-PS interactions are not accurately described by the tested models.

Our results pave the way for improving the PS lipid parameters for MD simulations
by offering the set of experimental data for the quality measurement, by
pinpointing problems areas in the models and suggesting directions for the corrections.
Improvements using the electronic continuum correction is already in progress \url{https://github.com/jmelcr/ecc_lipids},
following the recent work for PC lipids \cite{melcr18}.

>>>>

\section{Conclusions}
Lipids with PS headgroups and their interactions with ions
play an important role in lipid mediated signaling
processes \cite{leventis10,yeung08}. Recent studies using molecular
dynamics simulations to interpret the various spectroscopic data
give contradictory results for the calcium binding details to PS
headgroups \cite{melcrova16,valentine18,hallock18}.  
Here, we used the headgroup C-H bond order parameters and open collaboration method to evaluate the quality
of headgroup structure and ion binding affinity to PS lipids
in available MD simulation force fields,
as previously done for PC lipids \cite{botan15,catte16}.

First, we complemented the available experimental data of PS
lipid headgroup order parameters \cite{browning80,roux90} by measuring the signs of the order parameters.
None of the available force fields tested using the NMRlipids open
collaboration was accurate enough to reproduce the PS
headgroup order parameters with the experimental accuracy, but
the best models suggested a characteristic rigid conformation for the
carboxyl group in the serine headgroup. Comparison to the previously
measured headgroup order parameters from POPC:POPS (5:1) bilayers with different ion
concentrations \cite{roux90} showed that the tested MD simulation force fields
overestimate the cation binding affinity to the negatively charged bilayers
containing PS lipids with two exceptions. The apparently most realisitic monovalent ion binding
affinity to PS containing lipid bilayers was observed in the MacRog
simulations with potassium counterions and the CHARMM36 force field with recently introduced
NBfix correction for calcium \cite{kim16} underestimated the calcium binding affinity.
However, the experimentally measured trends of the PS headgroup order parameter response
to the bound cations and to the dilution of bilayer with zwitterionic PC lipids were not
qualitatively reproduced in any of the tested force fields, indicating that improvements in
the MD simulation force fields are necessary to study interactions between PS lipids and other biomolecules.
This is different to the previous results with PC lipids,
where the experimentally measured headgroup order parameter responses to the bound charge
were qualitative reproduced even thought the headgroup structures without ions were not
correct and the cation binding affinities were overestimated \cite{catte16}.

Our results pave the way for the development of better MD simulations force fields
for PS lipids. Using the headgroup order parameters, we were able to evaluate the
quality of various conformational ensembles in different force fields.
This can guide the development of force fields that would correctly reproduce the
conformations sampled by PS headgroups. The experimental dataset of headgroup order
parameters from POPC:POPS (5:1) mixture with different cation concentrations can
be used to improve cation binding details in the force fields, as recently demonstrated
for POPC using the electronic continuum correction \cite{melcr18}.
Similar study for POPS is progressed separately \cite{ECCpops}.
